A Lyapunov Analysis of Momentum Methods \\ in Optimization

### 2016-10-28 LyapDraft.tex
Update LyapDraft.tex

As mentioned in Section~\ref{Sec:StrongConv}, the dynamics~\eqref{Eq:StrongDyn} developed for the setting when $f$ is strongly convex can be viewed variationally, as the Euler-Lagrange equation for a different Lagrangian functional. %A quick check shows that it is the Euler-Lagrange equation for the following Lagrangian
%\begin{equation}\label{Eq:LagStrongConv}
%\mathcal{L}(x, \dot x, t) = \dot \beta_t e^{2\beta_t} \left(\frac{\mu}{2}\left\|\frac{\dot x}{\dot \beta_t}\right\|^2 - f(x)\right)
%\end{equation}
While outside the scope of this paper, the Lagrangian~\eqref{Eq:LagStrongConv} can be generalized to a non-Euclidean setting. In follow-up work, we plan to study this second Lagrangian family and its properties further. A natural question this research raise is whether the 
the Lyapunov analysis introduced in this paper can be extended other settings, such as accelerated coordinate ascent~\cite{Zeyuan16} and stochastic gradient descent, and whether a dynamical perspective can be developed in these settings.  %. Preliminary results suggests that this extension is indeed possible. %In followup work, we will present these extensions. 
%Of particular interest will be understanding the dynamical perspective these extensions potentially provide when randomness is introduced to~\eqref{Eq:EL} and~\eqref{Eq:StrongDyn}. %Furthermore, recently there has been several connections made between optimization and sampling schemes~\cite{ }

Earlier work by Drori and Teboulle~\cite{DroriTeboulle13}, Kim and Fessler~\cite{Kim2016}, Taylor \emph{et al}~\cite{Taylor2016}, and Lessard~\emph{et al}~\cite{Lessard14} have shown that optimization algorithms could be analyzed by solving convex programming problems.  In particular, Lessard~\emph{et al} show that Lyapunov-like potential functions called \emph{integral quadratic constraints} can be found by solving a constant-sized semidefinite programming problem. It would be interesting to see if these results can be adapted to directly search for Lyapunov functions like those studied in this paper.  This would provide a method to automate the analysis of new techniques, possibly moving beyond momentum methods to novel families of optimization techniques.
\section*{Acknowledgements}
We would like to give special thanks to Andre Wibisono for the many helpful discussion involving this paper. We would also like to thank Orianna Demassi for her very useful feedback and Stephen Tu who caught a small error in a previous version of this paper.
\bibliographystyle{plain}
\bibliography{refs.bib}

>>>>

As mentioned in Section~\ref{Sec:StrongConv}, the dynamics~\eqref{Eq:StrongDyn} developed for the setting when $f$ is strongly convex can be viewed variationally, as the Euler-Lagrange equation for a different Lagrangian functional. %A quick check shows that it is the Euler-Lagrange equation for the following Lagrangian
%\begin{equation}\label{Eq:LagStrongConv}
%\mathcal{L}(x, \dot x, t) = \dot \beta_t e^{2\beta_t} \left(\frac{\mu}{2}\left\|\frac{\dot x}{\dot \beta_t}\right\|^2 - f(x)\right)
%\end{equation}
While outside the scope of this paper, the Lagrangian~\eqref{Eq:LagStrongConv} can be generalized to a non-Euclidean setting. In follow-up work, we plan to study this second Lagrangian family and its properties further. A natural question is whether the 
the Lyapunov analysis introduced in this paper can be extended to other settings, such as accelerated coordinate ascent~\cite{Zeyuan16} and stochastic gradient descent, and whether a dynamical perspective can be developed in these settings.  %. Preliminary results suggests that this extension is indeed possible. %In followup work, we will present these extensions. 
%Of particular interest will be understanding the dynamical perspective these extensions potentially provide when randomness is introduced to~\eqref{Eq:EL} and~\eqref{Eq:StrongDyn}. %Furthermore, recently there has been several connections made between optimization and sampling schemes~\cite{ }

Earlier work by Drori and Teboulle~\cite{DroriTeboulle13}, Kim and Fessler~\cite{Kim2016}, Taylor \emph{et al}~\cite{Taylor2016}, and Lessard~\emph{et al}~\cite{Lessard14} have shown that optimization algorithms could be analyzed by solving convex programming problems.  In particular, Lessard~\emph{et al} show that Lyapunov-like potential functions called \emph{integral quadratic constraints} can be found by solving a constant-sized semidefinite programming problem. It would be interesting to see if these results can be adapted to directly search for Lyapunov functions like those studied in this paper.  This would provide a method to automate the analysis of new techniques, possibly moving beyond momentum methods to novel families of optimization techniques.

\section*{Acknowledgements}
We would like to give special thanks to Andre Wibisono for the many helpful discussion involving this paper. We would also like to thank Orianna Demassi for her very useful feedback and Stephen Tu who caught a small error in a previous version of this paper.
\bibliographystyle{plain}
\bibliography{refs.bib}
